---
title: "iris_markdown"
author: "Kevin Richard"
date: "2024-09-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(caret)
library(randomForest)
```

## Iris Flower Petal Dataset

```{r iris}
data(iris)
glimpse(iris)
```

## Including Plots


```{r plot1, echo=FALSE}
pairs(iris, col = iris$Species)
```

### The best correlation can be seen between the species and other features


```{r plot2, echo=FALSE}
iris_long <- pivot_longer(iris, cols = -Species, names_to = "Feature", values_to = "Value")
ggplot(iris_long,aes(x = Feature, y=Value)) +
  geom_boxplot(aes(fill=Species), outliers = FALSE)+
  scale_color_brewer(palette="Dark2")+
  theme_minimal()
```

## Descerning Factors
For each different Specie : 
Sepal.Length > Sepal.Width > Petal.Length > Petal.Width

Setosa -> has smaller Petal.Length,Petal.Width and Sepal.Length  than other species but has a larger Sepal.Width.

Versicolor and Virgincia -> have similar features but Virgincia is slighty bigger.


## RandomForsetModel for predicting Flower Specie

```{r RandomForsetModel, echo=FALSE}
set.seed(42)
trainIndex <- createDataPartition(iris$Species, p = 0.8, list = FALSE)
iris_train <- iris[trainIndex, ]
iris_test <- iris[-trainIndex, ]

rf_model <- randomForest(Species ~ ., data = iris_train, ntree = 1000, random_state = 42)

rf_predictions <- predict(rf_model, iris_test)
confusion_matrix <- confusionMatrix(rf_predictions, iris_test$Species)
print(confusion_matrix)

input_data <- data.frame(
  Sepal.Length=c(5.1,6.2,5.7),
  Sepal.Width=c(3.5,3.2,2.8),
  Petal.Length=c(1.4,4.5,4.5),
  Petal.Width=c(0.2,1.5,1.3)
)

print("Predicting for following input")
print(input_data)

output <- predict(rf_model, input_data)
print(output)
```
