---
title: "crickets_markdown"
author: "Kevin Richard"
date: "2024-08-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(modeldata)
```

## Cricket Dataset analysis

### Dataset Format
```{r glimpse}
glimpse(crickets)
```

### Species categories data count
```{r dataset, echo=FALSE}
ggplot(crickets, aes(x=species, fill=species))+
  geom_bar(width = 0.5)+
  labs(
    x = "species",
    y = "Count",
    title = "Data count of Species",
    fill = "Species",
  )+
  scale_color_brewer(palette="Dark2")+
  theme_minimal()
```

## Basic Plots of categories to quantitative data cloumn

```{r plots, echo=FALSE}
ggplot(crickets, aes(x=temp, fill=species))+
  geom_histogram(binwidth  = 0.5)+
  labs(
    x = "Temp",
    y = "Count",
    title = "Temp count",
    fill = "Species",
  )+
  facet_wrap(~species, ncol=1)+
  scale_color_brewer(palette="Dark2")+
  theme_minimal()

ggplot(crickets, aes(x=rate, fill=species))+
  geom_histogram()+
  labs(
    x = "Chirp Rate",
    y = "Count",
    title = "Chirp Rate count",
    fill = "Species",
  )+
  scale_color_brewer(palette="Dark2")+
  theme_minimal()
```

## Temperature to Chirp rate relation
```{r relations 1, echo=FALSE}
ggplot(crickets, aes(x=temp, y=rate, color=species)) +
  geom_boxplot() +
  labs(
    x = "Temperature",
    y = "Chirp Rate",
    title = "Chirp Rate to Temperature",
    color = "Species"
  )+
  scale_color_brewer(palette="Dark2")+
  theme_minimal()
```

### It can be noted that O.e spicies have a higher chirp rate than O.n

```{r relations 2, echo=FALSE}
ggplot(crickets, aes(x=temp, y=rate, color=species))+
  geom_point(size = 1.5, alpha = 1, shape = "circle")+
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.5)+
  labs(
    x = "Temperature",
    y = "Chirp Rate",
    title = "Chirp Rate to Temperature",
    color = "Species"
  )+
  scale_color_brewer(palette="Dark2")+
  theme_minimal()
```

### From above it can be noted that :-
### 1) The Chirp rate increases as the temperature raises irrespective of species
### 2) Based on the available data: O.e has higher chirp rate than O.e
